Constraints on Assembly Bias from Galaxy Clustering.

### 2016-06-08 ms.tex
Made number of small corrections to intro

For more than a decade, halo occupation modeling has been used to
interpret large-scale structure measurements and exploit these
measurements to constrain galaxy formation models and cosmology
\citep[e.g.,][]{yang03,tinker05,zehavi05a,
  porciani06,vdBosch07,Zheng07,conroy_wechsler09,yang09b,zehavi_etal11,guo_etal11b,
  wake_etal11,yang11a,yang12,leauthaud_etal12,rod_puebla12,tinker_etal13,cacciato_etal13,
  more_etal13,guo_etal14,zu_mandelbaum15b}. The key assumptions
underlying halo occupation modeling are: (1) all galaxies reside in
dark matter halos that are biased tracers of the density field; and
(2) galaxies occupy halos as a function of halo masses only. It is now
well known that halo bias depends upon halo properties other than mass
\citep[e.g.][]{gao_etal05,wechsler06,gao_white07,zentner07,dalal_etal08,lacerna11},
an effect called halo assembly bias. If galaxies occupy halos as a
function of properties other than halo mass, then standard halo
occupation methods will be subject to a systematic error due to galaxy
assembly bias. Several of us have previously shown that this error can
be significant in an analysis of galaxy clustering and can bias
inferences about many aspects of galaxy evolution
\citep{zentner_etal14}. Consequently, we have developed halo
occupation models that enable galaxies to occupy halos in a manner
that depends upon several halo properties \citep{hearin_etal16}.  See
also \citet{Yao_etalXX} for a similar approach. In this paper, we
revisit the interpretation of luminosity-dependent galaxy clustering
in the Sloan Digital Sky Survey (SDSS) Data Release 7 (DR7) data,
analyzed previously by \citet{zehavi_etal11}, in the context both
standard halo occupation models and these new models.

@@

Since, at fixed halo mass, $\vmax$ is correlated with halo
concentration, and thus with halo assembly time, the resulting
occupation models have assembly bias build in
\citep[see][]{zentner_etal14}; i.e., at fixed halo mass, galaxy
properties will be correlated with halo assembly
time. \citet{lehmann_etal15} took this one step further, and explored
SHAM models in which the halo property varies between $M$ and $\vmax$
via a continuously-valued parameter. Using SDSS clustering
measurements to constrain their model, they conclude that galaxy
luminosity reveals a significant dependence on halo
concentration. Since the latter is strongly correlated with halo
formation time \citep[e.g.,][]{wechsler02,zhao_etal09,ludlow_etal13},
their conclusions argue for non-zero assembly bias in the galaxy
population.

Additional support for the presence of assembly bias has come from a
variety of studies that have shown, or suggested, that the large-scale
environment (i.e., clustering strength) of dark matter haloes {\it of
  fixed mass} is correlated with the star formation rate of their
central
galaxies\citep[][]{yang_etal06a,blanton_berlind07,wang_etal08,tinker_etal12,wang_etal13,HW13a,hearin_etal13b,hearin_etal14,watson_etal14}.
However, \citet{lin_mandelbaum_etal15} claim that most of the these
differences in clustering strength can be explained as due to
differences in halo mass or contamination by satellite galaxies, and
therefore argue that there is no significant evidence for assembly
bias in galaxy-scale halos \citep[see also][]{paranjape15}.  On the
other hand, \citet{miyatake_etal15} and \citet{more_etal16} recently
presented strong evidence for the presence of assembly bias in massive
clusters; in particular, they find that clusters with a more
concentrated distribution of satellite galaxies are less strongly
clustered than their equal-mass counterparts whose satellite galaxies
are less centrally concentrated. To summarize, while there is strong
evidence for the presence of assembly bias on the scale of clusters,
there is not concensus yet as to whether assembly bias plays also a
role in less massive haloes.

>>>>

For more than a decade, halo occupation modeling has been used to
interpret large-scale structure measurements and exploit these
measurements to constrain galaxy formation models and cosmology
\citep[e.g.,][]{yang03,tinker05,zehavi05a,
  porciani06,vdBosch07,Zheng07,conroy_wechsler09,yang09b,zehavi_etal11,guo_etal11b,
  wake_etal11,yang11a,yang12,leauthaud_etal12,rod_puebla12,tinker_etal13,cacciato_etal13,
  more_etal13,guo_etal14,zu_mandelbaum15b}. The key assumptions
underlying halo occupation modeling are: (1) all galaxies reside in
dark matter halos that are biased tracers of the density field; and
(2) galaxies occupy halos as a function of halo masses only. It is now
well known that halo bias depends upon halo properties other than mass
\citep[e.g.][]{gao_etal05,wechsler06,gao_white07,zentner07,dalal_etal08,lacerna11},
an effect called halo assembly bias. If galaxies occupy halos as a
function of properties other than halo mass, then standard halo
occupation methods will be subject to a systematic error due to galaxy
assembly bias. Several of us have previously shown that this error can
be significant in an analysis of galaxy clustering and can bias
inferences about many aspects of galaxy evolution
\citep{zentner_etal14}. Consequently, we have developed halo
occupation models that enable galaxies to occupy halos in a manner
that depends upon several halo properties \citep{hearin_etal16}.  In
this paper, we revisit the interpretation of luminosity-dependent
galaxy clustering in the Sloan Digital Sky Survey (SDSS) Data Release
7 (DR7) data, analyzed previously by \citet{zehavi_etal11}, in the
context both standard halo occupation models and these new models.

@@

Since, at fixed halo mass, $\vmax$ is correlated with halo
concentration, and thus with halo assembly time, the resulting
occupation models have assembly bias build in
\citep[see][]{zentner_etal14}; i.e., at fixed halo mass, galaxy
properties will be correlated with halo assembly
time. \citet{lehmann_etal15} took this one step further, and explored
SHAM models in which the halo property varies between $M$ and $\vmax$
via a continuously-valued parameter. Using SDSS clustering
measurements to constrain their model, they conclude that at fixed
halo mass galaxy luminosity reveals a significant dependence on halo
concentration. Since the latter is strongly correlated with halo
formation time \citep[e.g.,][]{wechsler02,zhao_etal09,ludlow_etal13},
their conclusions argue for non-zero assembly bias in the galaxy
population.

Additional support for the presence of assembly bias has come from a
variety of studies that have shown, or suggested, that the large-scale
environment (i.e., clustering strength) of dark matter haloes {\it of
  fixed mass} is correlated with the star formation rate of their
central
galaxies\citep[][]{yang_etal06a,blanton_berlind07,wang_etal08,tinker_etal12,wang_etal13,HW13a,hearin_etal13b,hearin_etal14,watson_etal14}.
However, \citet{lin_mandelbaum_etal15} claim that most of these
differences in clustering strength can be explained as due to
differences in halo mass or contamination by satellite galaxies, and
therefore argue that there is no significant evidence for assembly
bias in galaxy-scale halos \citep[see also][]{paranjape15}.  On the
other hand, \citet{miyatake_etal16} and \citet{more_etal16} recently
presented strong evidence for the presence of assembly bias in massive
clusters; in particular, they find that clusters with a more
concentrated distribution of satellite galaxies are less strongly
clustered than their equal-mass counterparts whose satellite galaxies
are less centrally concentrated. To summarize, while there is strong
evidence for the presence of assembly bias on the scale of clusters,
there is not concensus yet as to whether assembly bias plays also a
role in less massive haloes.
